A Case Study in Preserving a High Energy Physics Application

### 2014-01-29 paper.tex
Cleaned up introduction.

\section{Overview of Application}

Matthias, one graduate student from Physics Department of University of Notre
Dame, finished one interesting tau analysis. Others want to understand how the tau
analysis works and repeat it.

\subsection{Background}

Within the ongoing investigation of the Higgs boson at the CMS
detector, part of the LHC at CERN, the Higgs production in association
with two top quarks allows to measure the Higgs coupling strength to
top quarks.  As the Higgs boson is to short-lived to be detected
itself, it has to be reconstructed from its decay products.  The
analysis discussed in this section focuses on the Higgs decaying to
two tau leptons, which are themselves reconstructed from particle
showers.  Within the analysis, events are identified that show a
signature of decay products compatible with both hadronic tau and top
decays.  Properties of such events are used to distinguish signal-like
(matching simulated Higgs decays) and background-like events and are
also used in further statistical analysis.

The CMS collaboration provides analysis end-users with a pre-processed
and reduced data format, AOD, containing information for events, i.e.,
proton-proton collisions with a signature of interest, in the form of
reconstructed particles.  This format is based on the RAW output of
the CMS detector readout electronics and reconstructed world-wide.
Both real and simulated data are available for examination.

@@

% FWIW, units are TB, GB, not TiB, GiB!

@@

Understanding all the remote and local dependencies of one program, one self-contained package can be built from scratch by copying all the dependencies directly using coarse-grained method,
which means copying the whole local filesystem, HDFS, PanFS, together with each complete remote repository without selection, into the new package.
However, the whole OS and local filesystem do not exist only for this special program, and contain a lot of files and software unrelevant to this special program. 
The size of HDFS, PanFS, remote CMSSW repository all reaches TB level. 
This solution is easy to implement, but time-consuming and space-consuming.
Through observation, we noticed that the size of really used data and software is greatly less than the original total size. 
Therefore, one fine-grained toolkit generating the target package is necessary.

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=.8\columnwidth]{data-code-size.eps}
\caption{Inputs to Tau Analysis}
\label{fig:data-code-size}
\end{figure}

\begin{table}[t]
    \centering
    \small
    \begin{tabular}{|l|l|r|r|r|}
        \hline
        \bf Name & \bf Location & \bf Total & \bf Named & \bf Used \\ 
        \hline
        SBEAN data    & hdfs & 33TB & 43.3GB & 20GB \\ \hline
        CMSSW binaries     & local & 1.16TB & 449MB & 5.2MB\\ \hline
        Tau source       & git & 51MB & 51MB & 51MB \\ \hline
        PyYAML binaries    & http & 52MB & 51MB & 51MB \\ \hline
        .h file       & http & 41KB & 41KB & 41KB \\ \hline
        Misc commands & hpc fs & 206TB & ??? & 119MB \\ \hline
        Linux commands & local fs & ??? & ??? & 88.4MB \\ \hline     
        Total          &          & ??? & ??? & ??? \\ \hline
    \end{tabular}
    \caption{Data and Code Used by Tau Analysis}
    \label{table:size-original-real}
\end{table}

\section{Overview of Tau Analysis}

Within the ongoing investigation of the Higgs boson at the CMS
detector, part of the LHC at CERN, the Higgs production in association
with two top quarks allows to measure the Higgs coupling strength to
top quarks.  As the Higgs boson is to short-lived to be detected
itself, it has to be reconstructed from its decay products.

The application which is the study of this paper is called \emph{TauAnalysis}
It searches for cases where the the Higgs boson decays to two tau leptons.
The leptons are not observed directly, but by the particle showers
that they generate.  So, the analysis must search for detector
events that show a signature of decay products compatible with both hadronic tau and top decays.  Properties of such events are used to distinguish
the events of interest (Higgs decays) from all other events and
are also used in further statistical analysis.

Figure~\ref{fig:data-code-size} shows that both the code and data
that form \emph{TauAnalysis} are drawn from large repositories through
multiple steps of reduction.

{\bf Data Sources.}
The CMS collaboration provides analysis end-users with a pre-processed
and reduced data format, AOD, containing information for events, i.e.,
proton-proton collisions with a signature of interest, in the form of
reconstructed particles.  This format is based on the RAW output of
the CMS detector readout electronics and reconstructed world-wide.
Both real and simulated data are available for examination.

@@

{\bf Code Sources.} Like many scientific codes, the central algorithm
of \emph{TauAnalysis} is expressed in a relatively small amount of
custom code developed by the primary author.  But, the code cannot
run at all without making use of an enormous collection of software
dependencies.  Some of these dependencies are standard to operating
systems worldwide, some are standardized across the entire high-energy
physics field, some are particular to small collaborative groups,
and a few are very specific to a single researcher.

The largest of these repositories is the CMS Software Distribution (CMSSW),
a carefully-curated selection of software packages which is distributed
in binary form.  Historically, CMSSW was downloaded and installed on shared
filesystems within HPC centers.  In recent years, distribution has moved to
an on-demand delivery system known as CVMFS~\cite{cvmfs}.  The content
of CMSSW is managed very carefully by a centralized team whose main goal
is to ensure that the current version of the software operates correctly
on the operating systems and architectures currently in use.  However,
preservation per se is not an objective of the group, and so there is
no guarantee that old versions of CMSSW operate in new environments,
or vice versa.

\begin{figure*}
\centering
\includegraphics[width=1.6\columnwidth]{version-evolution.eps}
\caption{Version Evolution}
\label{fig:version-evolution}
\end{figure*}

\emph{TauAnalysis} was provided to us in the form of an
email which described, in prose, how to obtain the source,
build the program, and run it correctly on one specific
machine at Notre Dame, with no particularly guarantee that
it will run anywhere else in the world.
Although this starting point may seem extreme, it is
perfectly natural for collaborators to share configurations
with each other in this form, and to rely on the presence
of a working environment with appropriate dependencies already
installed.

The authors played the role of curators, whose job it is to prepare the application for permanent archival by determining and packaging the dependencies.

@@
